---
title: "607 HW 9"
author: "Carlisle Ferguson"
date: "4/10/2021"
output: 
  html_document:
      theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Web APIs
Assignment Goal: Choose a NYT API, build an interface in R to read in JSON data, and transform it into an R dataframe. 

## Loading Libraries
```{r}
library(jsonlite)
library(tidyverse)
library(kableExtra)
```
## Reading and Transforming JSON Data
The `jsonlite` library has a function called `fromJSON` that can be used to read a JSON file from a URL. To read a JSON file off a NYT API, users also need a key, which can be obtained by following instructions on NYT's website (link [here](https://developer.nytimes.com/get-started)). For this assignment, I chose to use the book API, and used the example query of Stephen King's bestseller reviews. I also modified the query to instead search for reviews of Michelle Obama's bestseller. The data frames are displayed below.

```{r}
data_ex <- fromJSON("https://api.nytimes.com/svc/books/v3/reviews.json?author=Stephen+King&api-key=grSS1CvDQBtl4ZyUk1ZSkASsTbGR5mru", flatten = TRUE) %>% data.frame()
data_ex %>%
  kbl(caption = "Reviews of Stephen King's Bestsellers") %>%
  kable_styling(bootstrap_options = "striped")
data <- fromJSON("https://api.nytimes.com/svc/books/v3/reviews.json?author=Michelle+Obama&api-key=grSS1CvDQBtl4ZyUk1ZSkASsTbGR5mru", flatten = TRUE) %>% data.frame()
data %>%
  kbl(caption = "Review of Michelle Obama's Best Seller") %>%
  kable_styling(bootstrap_options = "striped")
```
## Conclusion
This was relatively straightforward with the use of the `jsonlite` library, and showed how one can easily access a large amount of data.


```


